---
title: "R Code"
---

## Writing better code

We can improve our scripts by writing code which is:

1. Readable

- Use a consistent style guide (like the tidyverse one)
- Be clear, concise, and consistent when naming objects, datasets, and functions
- Break your scripts up into sections (Ctrl+Shift+R)

2. Reusable

- Write functions to perform sets of operations
- Use iteration (for loops, or `map()`) 
- Consider creating packages for data and code used in multiple places
- Break up long scripts into separate files (cleaning, manipulating, plotting)

3. Documented

- Use comments to explain why you did something the way you did 
- Include a README with each project to provide important context and background

Create a new script in RStudio (Ctrl+Shift+N) and save it in the `scripts` folder.

Load packages we need for the workshop.

```{r}
library(readr)
library(readxl)
library(dplyr)
library(ggplot2)
library(purrr)
library(renv)
```

Let's take a common example of some data analysis and apply the above principles when writing the code.

First, let's read in some data split into different files by month. Instead of naming each file, we can use `readr::read_csv()` to read them all in at once, bind them together into a single dataframe, and add a new column containing the filename of the data source. 

```{r}
files <- list.files(path = "data_raw", pattern = "\\.csv", full.names = TRUE)

data <- read_csv(files, id = "source")
```


